---
title: "How to use Excyte"
author: "Maxime Meylan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction

Excyte is a pipeline that allows exhaustive exploration of cytometry data. Excyte first preprocesses compensated fcs data, then performs unsuppervised clustering of selected events with PhenoGraph (Jacob H. Levine et.al; Cell, 2015) and dimensionality reduction via Umap (Leland McInnes et.al; arXiv:1802.03426). This pipeline outputs proportions of identified clusters for each input sample as well as intuitive visualizations.


## Get started 

In R, install Excyte via GitHub with the command 

``` 
library(devtools)
install_github("maxmeyl/Excyte")
```


## First run 

To run the Excyte pipeline, input the directory containing compensated fcs files. Downsampling can be used to speed up the computation time, here 1000 events of each fcs are randomly sampled. Finally, we will run Excyte on avalaible channels.


```{r} 
library(excyte)
library(gridExtra)

#load sample data 
fcs.loc <- system.file("extdata",package="flowCore")
file.location <- paste(fcs.loc, dir(fcs.loc), sep="/")

excyte_object <- run_excyte(fcs_dir = file.location[1:3],downsampling = 2000,channels = "all")
```
## Results of the pipeline

`excyte_object` contains the Umap and clustering results \

`excyte_object$processed_fcs_obj` contains the normalized intensities of each event \

`excyte_object$pheno_obj`contains the output from PhenoGraph clustering \

`excyte_object$excyte_res$umap_obj` contains the ouput from the Umap dimensionality reduction, including the 2D coordinates of each event.\

The percentage of clusters among each samples can be displayed by :

```{r, echo=FALSE, results='asis'}
knitr::kable(head(excyte_object$phenograph_obj$phenograph_percentage))
```

## Visualizations 

Excyte propose several visualization than can be computed with the command
```{r}
excyte_plots <- plot_excyte(excyte_obj = excyte_object,alpha = 0.7)
```

The visualizations can be displayed by querying the result object 

Heatmap presenting the average intensity profile of each cluster
```{r , fig.width = 4,fig.height=4,dpi=100} 
excyte_plots$heatmap
```

```{r , fig.width = 8,fig.height=7,dpi=100} 
grid.arrange(grobs=excyte_plots$umap_channels[1:4],ncol=2)
```

```{r , fig.width = 8,fig.height=7,dpi=100} 
grid.arrange(grobs=excyte_plots$ridges[1:4],ncol=2)
```

## Re-run excyte on selected clusters 

The pipeline can be re-run on a subset of the cluster to better characterize interesting populations 

```{r,fig.width = 8,fig.height=7,dpi=100}
rerun_object <- rerun_excyte(excyte_obj = excyte_object,clusters_id = c("c21","c3","c9","c11","12"),k=30)
rerun_plots <- plot_excyte(excyte_obj = rerun_object,alpha = 1)
grid.arrange(grobs=excyte_plots$umap_channels[1:4],ncol=2)

```


The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
