---
title: "How to use Excyte"
author: "Maxime Meylan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Description

Excyte is a pipeline that allows exhaustive exploration of cytometry data. Excyte first preprocesses compensated fcs data, then performs unsuppervised clustering of selected events with PhenoGraph (Jacob H. Levine et.al; Cell, 2015) and dimensionality reduction via Umap (Leland McInnes et.al; arXiv:1802.03426). This pipeline outputs proportions of identified clusters for each input sample as well as intuitive visualizations.


## Get started 

In R, install Excyte via GitHub with the command 

``` 
install.packages("devtools")
library(devtools)
install_github("maximemeylan/Excyte")
```


## First run 

To run the Excyte pipeline, input the directory containing compensated fcs files. Downsampling can be used to speed up the computation time, here 1000 events of each fcs are randomly sampled. Finally, we will run Excyte on avalaible channels.


```{r} 
library(gridExtra)
library(excyte)

#load sample data 
fcs.loc <- system.file("extdata",package="excyte")
file.location <- paste(fcs.loc, dir(fcs.loc), sep="/")

#for reproducible output
set.seed(8)

#run the pipeline
excyte_object <- run_excyte(fcs_dir = file.location[1:3],
                            downsampling = 3000,
                            channels = "with_desc",
                            downsampling_umap = 1000)
```
## Results of the pipeline

`excyte_object` contains the result objects computed by the pipeline\

`excyte_object$processed_fcs_obj` contains the normalized intensities of each event \

`excyte_object$pheno_obj`contains the object from PhenoGraph clustering \

`excyte_object$excyte_res$umap_obj` contains the ouput from the Umap, including the 2D coordinates of each event.\

The percentage of clusters among each samples can be displayed by :
`excyte_object$phenograph_obj$phenograph_percentage`

```{r, echo=FALSE, results='asis'}
knitr::kable(excyte_object$phenograph_obj$phenograph_percentage[1:3,1:5])
```

## Visualizations 

Excyte propose several visualization than can be computed with the command
```{r}
excyte_plots <- plot_excyte(excyte_obj = excyte_object,alpha = 0.7)
```

The visualizations can be displayed by querying the result object 

### Heatmap presenting the average intensity profile of each cluster

```{r , fig.width = 8,fig.height=7,dpi=100} 
excyte_plots$heatmap
```

### Umap of event intensities for selected channels
```{r , fig.width = 8,fig.height=7,dpi=100} 
grid.arrange(grobs=excyte_plots$umap_channels[c(9,3,4,1,2,5)],ncol=2)
```

### Umap of cluster membership
```{r , fig.width = 8,fig.height=7,dpi=100} 
excyte_plots$umap_phenograph
```


### Distribution of event intensities according to cluster membership
```{r , fig.width = 8,fig.height=7,dpi=100} 
grid.arrange(grobs=excyte_plots$ridges_clusters[c(9,3,4,1,2,5)],ncol=2)
```

## Re-run excyte on selected clusters 

The pipeline can be re-run on a subset of the cluster to better characterize interesting populations 
Here we focus on CD3, CD4 high clusters, and investigate memory and activation markers

```{r,fig.width = 8,fig.height=7,dpi=100}
set.seed(8)
excyte_object$processed_fcs_obj$all_channels
markers <- excyte_object$processed_fcs_obj$all_channels[c(c(11,7,12,13,8,14)),1]

rerun_object <- rerun_excyte(excyte_obj = excyte_object,
                             clusters_id = c("C_19","C_12","C_18","C_13","C_05","C_08"), 
                             channels = markers,)
rerun_plots <- plot_excyte(excyte_obj = rerun_object,alpha = 1)
grid.arrange(grobs=rerun_plots$umap_channels,ncol=2)
```
### Umap of cluster membership
```{r , fig.width = 8,fig.height=7,dpi=100} 
rerun_plots$umap_phenograph
```

